---
title: CohortOperationsShinyApp
author: Javier Gracia-Tabuenca
output:
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

If you are only interested into running the app already installed inside SandBox, you can jump to [Running in Sandbox](#running-in-sandbox)

## Intro

The goal of CohortOperationsShinyApp is to provide a graphical interface to create new cohorts based on cohorts imported from Atlas or from a text file. 

A cohort in this tool follows the epidemiological definition: "A cohort is a set of persons who satisfy one or more inclusion criteria for a duration of time" [[1]](https://ohdsi.github.io/TheBookOfOhdsi/Cohorts.html#what-is-a-cohort). 

This web app is a graphical interface to other R-packages which provide the underlying functionality. 
For programming use cases check these: 

- [FinnGen/FinnGenTableTypes](https://github.com/FinnGen/FinnGenTableTypes): Provides tools to create and operate cohortData tables. 
- [FinnGen/CDMTools](https://github.com/FinnGen/CDMTools): Provides connection to Atlas and Atlas' CDM databases. 

This project follows [Golem](https://engineering-shiny.org/golem.html) philosophy for the development of Shiny Apps, including modularisation, unit testing, shiny server testing, yalm configuration, and dockerisation. Moreover, this projects uses [renv](https://rstudio.github.io/renv/articles/renv.html) for package dependency management. 

## Development 
Before any of the following steps you need a github token to access the code in the private repositories. Set the environmental variable `GITHUB_PAT` with a token generated in [github.com/settings/tokens](https://github.com/settings/tokens). Token should include following permissions: "gist, repo, user, workflow". Following code helps you to do that from R: 

```{r, eval=FALSE}
# generate tocken 
usethis::create_github_token()

# copy and paste in 
Sys.setenv(GITHUB_PAT="<paste_token>")
```

### Development in laptop 
For development you can clone this repository and use `renv::` to install all the dependent packages. 

`renv::` automatically installs its self at the opening of the project. Then, run `renv::restore()` to install the dependent packages. 

### Development in SandBox 
SandBox has not connection to the internet. 
However, dependencies cant be build in a temporal IVM with connection to the internet, zipped, and copy into SandBox IVM. 
(note: the temporal IVM should have the same operating system and R version)

Follow `renv::` instruction for a such situation: 

> Clone the package on a computer with internet connection. 
> Specify the folder where your packages are stored by setting the RENV_PATHS_CACHE location (run `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))`).
> Then run `renv::restore()` in the console. 
> Manually move the study package to the environment without internet (this now includes all required R packages), activate the current project with `renv::activate()` and again run `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))` followed by `renv::restore()` in the console.


### Configure development enviroment
Configuration for your development environment can be set in `inst/golem-config.yml`. 
And selected on run time seting the envar `GOLEM_CONFIG_ACTIVE` (see: [golem-config](https://engineering-shiny.org/golem.html?q=GOLEM_CONFIG_ACTIVE#golem-config))

Currently this file includes three environments: 

- `no_connection`: To work with no connection to an Atlas instance. 
- `atlas-development`: To work with Atlas installed in an ivm in the atlas-development project in GCP. 
This environment needs the following additional yalm variables to configure [FinnGen/CDMTools](https://github.com/FinnGen/CDMTools).   - `CDMTOOLS_dbms` = "bigquery-dbi"
  - `GCP_PROJECT_ID` = "atlas-development-270609" 
  - `GCP_BILLING_PROJECT_ID` = "atlas-development-270609"
  - `CDMTOOLS_webapi_url` = "http://localhost/WebAPI"
  - `CDMTOOLS_CDM_source_key_test` = "dummy_df6v2_1k_13_FinnGen_omop_bq"
  - `GCP_SERVICE_KEY`: path to the GCP key to access the BQ in atlas-development
  
- `SandBox`: To work in SandBox.
SandBox must have set the following environmental variable `BUCKET_SandBox_IVM` with bigquery's billing project name ending in "_ivm" (eg. "fg-production-SandBox-4_ivm"). 
This environment  also needs the following yalm variables: 
  - `CDMTOOLS_dbms` = "bigquery-dbi"
  - `GCP_PROJECT_ID` = "FinnGen-production-library"
  - `CDMTOOLS_webapi_url` = "https://ohdsi-webapi.app.FinnGen.fi/WebAPI"
  - `CDMTOOLS_CDM_source_key_test` = "FinnGen_CDM_R7"

### Run in development
Open R in the package folder, or open project in Rstudio. 

```{r, eval=FALSE}
# load package
devtools::load_all(".")

# set configuration 
Sys.setenv(GOLEM_CONFIG_ACTIVE="<config_tag_in_golem-config.yml>")
# if this is not set, default configuration is no_connection environment
# Rstudio in SandBox is not reading the system environmental variables, force the envar as 
# Sys.setenv(BUCKET_SandBox_IVM="fg-production-SandBox-<n SandBox>_ivm")

# run shiny app
run_app()
)
```


## Deployment 

### Build docker image
Docker image can be build from scratch or, to save time, it can be built using the pre-compiled dependencies built in  above section "Development in SandBox". 

Both methods use the same command:

```{bash, eval=FALSE}
cd <CohortOperationsShinyApp>
sudo docker build -t cohort_operations_shiny_app --build-arg GITHUB_PAT=<paste_PAT_token> .
```

If `renv::restore()` was run with option `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))`. The cache is copied into the docker image during building. If so, you mush keep updated the cache if it changes during development. 

Alternatively, you can run `renv::restore()` or erase "./renv/cache" folder. In this case, all packages will be downloaded and install during the building process.  

### Load image into SandBox 
Built image can be moved into SandBox in two ways: (A) zipped and upload, or (B) through GCP Container Register. 

#### (A) Download and upload
Docker image can be save with: 

```{bash, eval=FALSE}
docker save --output cohort_operations_shiny_app.tar cohort_operations_shiny_app
```

Downloaded. For example using :

```{bash, eval=FALSE}
python3 -m http.server 8888
```

Uploaded and loaded into sanxbox:

```{bash, eval=FALSE}
docker load --input cohort_operations_shiny_app.tar
```

#### (B) Push and pull from GCP Container Register
Make sure you are [running docker without sudo](https://github.com/sindresorhus/guides/blob/main/docker-without-sudo.md). 

Authenticate with application-default login and configure docker. Use your FinnGen account
```{bash, eval=FALSE}
gcloud auth login
#gcloud auth application-default login #QUESTION: this was not working, is it ok the above way ??
gcloud auth configure-docker
# you can heck that credentials exist.
#cat ~/.config/gcloud/application_default_credentials.json
```

Tag the image.
```{bash, eval=FALSE}
docker tag cohort_operations_shiny_app eu.gcr.io/atlas-development-270609/cohort_operations_shiny_app:<version_tag>
```

Push newly tagged image to destination.
```{bash, eval=FALSE}
docker push eu.gcr.io/atlas-development-270609/cohort_operations_shiny_app:<version_tag>
```

Revoke credentials.
```{bash, eval=FALSE}
gcloud auth revoke
```

([other help link](https://cloud.google.com/container-registry/docs/advanced-authentication))

Request to humgen service desk to copy the image into SandBox container registry. 

Open a terminal inside SandBox. Pull the image :
```{bash, eval=FALSE}
docker pull eu.gcr.io/finngen-sandbox-v3-containers/cohort_operations_shiny_app:<version_tag>
```

## Running in SandBox 
Assuming the docker image is already deployed in your system. 

Find the image id: 
```{bash, eval=FALSE}
docker images
```

Run: 
```{bash, eval=FALSE}
docker run -p 8888:8888 -e BUCKET_SandBox_IVM=$BUCKET_SandBox_IVM <docker_image_id>
```

> Running image needs to tunnel the port (-p) and set envar `BUCKET_SandBox_IVM` (-e)
>
> TEMP: at the moment BUCKET_SandBox_IVM needs to be passed, bcs the docker is not getting the envars from the main. 
>
> By default, the docker image uses "production" configuration (see [inst/golem-config.yml](inst/golem-config.yml)). This can be over written by setting the envar `GOLEM_CONFIG_ACTIVE`. For example, `GOLEM_CONFIG_ACTIVE=default` runs the app without connection to Atlas. 

Browse to shiny server: 

Now, shiny server should be up and running. Connect to it with a web browser on `http://0.0.0.0:8888`. 










