---
title: CohortOperationsShinyApp
author: Javier Gracia-Tabuenca
output:
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Intro

The goal of CohortOperationsShinyApp is to provide a graphical interface to create new cohorts based on cohorts imported from Atlas or from a text file. 

A cohort in this tool follows the epidemiological definition: "A cohort is a set of persons who satisfy one or more inclusion criteria for a duration of time" [[1]](https://ohdsi.github.io/TheBookOfOhdsi/Cohorts.html#what-is-a-cohort). 

This web app is a graphical interface to other R-packages which provide the underlying functionality. 
For programming use cases check these: 

- [FINNGEN/FinnGenTableTypes](https://github.com/FINNGEN/FinnGenTableTypes): Provides tools to create and operate cohortData tables. 
- [FINNGEN/CDMTools](https://github.com/FINNGEN/CDMTools): Provides connection to Atlas and Atlas' CDM databases. 

This project follows [Golem](https://engineering-shiny.org/golem.html) philosophy for the development of Shiny Apps, including modularisation, unit testing, shiny server testing, yalm configuration, and dockerisation. Moreover, this projects uses [renv](https://rstudio.github.io/renv/articles/renv.html) for package dependency management. 

## Development 
Before any of the following steps you need a github token to access the code in the private repositories. Set the environmental variable `GITHUB_PAT` with a token generated in [github.com/settings/tokens](https://github.com/settings/tokens). Token should include following permissions: "gist, repo, user, workflow". Following code helps you to do that from R: 

```{r, eval=FALSE}
# generate tocken 
usethis::create_github_token()

# copy and paste in 
Sys.setenv(GITHUB_PAT="<paste_token>")
```

### Development in laptop 
For development you can clone this repository and use `renv::` to install all the dependent packages. 

`renv::` automatically installs its self at the opening of the project. Then, run `renv::restore()` to install the dependent packages. 

### Development in Sandbox 
Sandbox has not connection to the internet. 
However, dependencies cant be build in a temporal IVM with connection to the internet, zipped, and copy into Sandbox IVM. 
(note: the temporal IVM should have the same operating system and R version)

Follow `renv::` instruction for a such situation: 

> Clone the package on a computer with internet connection. 
> Specify the folder where your packages are stored by setting the RENV_PATHS_CACHE location (run `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))`).
> Then run `renv::restore()` in the console. 
> Manually move the study package to the environment without internet (this now includes all required R packages), activate the current project with `renv::activate()` and again run `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))` followed by `renv::restore()` in the console.


### Configure development enviroment
Configuration for your development environment can be set in `inst/golem-config.yml`. 
And selected on run time seting the envar `GOLEM_CONFIG_ACTIVE` (see: [golem-config](https://engineering-shiny.org/golem.html?q=GOLEM_CONFIG_ACTIVE#golem-config))

Currently this file includes three environments: 

- `no_connection`: To work with no connection to an Atlas instance. 
- `atlas-development`: To work with Atlas installed in an ivm in the atlas-development project in GCP. 
This environment needs the following additional yalm variables to configure [FINNGEN/CDMTools](https://github.com/FINNGEN/CDMTools).   - `CDMTOOLS_dbms` = "bigquery-dbi"
  - `GCP_PROJECT_ID` = "atlas-development-270609" 
  - `GCP_BILLING_PROJECT_ID` = "atlas-development-270609"
  - `CDMTOOLS_webapi_url` = "http://localhost/WebAPI"
  - `CDMTOOLS_CDM_source_key_test` = "dummy_df6v2_1k_13_finngen_omop_bq"
  - `GCP_SERVICE_KEY`: path to the GCP key to access the BQ in atlas-development
  
- `sandbox`: To work in sandbox.
Sandbox must have set the following environmental variable `BUCKET_SANDBOX_IVM` with bigquery's billing project name ending in "_ivm" (eg. "fg-production-sandbox-4_ivm"). 
This environment  also needs the following yalm variables: 
  - `CDMTOOLS_dbms` = "bigquery-dbi"
  - `GCP_PROJECT_ID` = "finngen-production-library"
  - `CDMTOOLS_webapi_url` = "https://ohdsi-webapi.app.finngen.fi/WebAPI"
  - `CDMTOOLS_CDM_source_key_test` = "FINNGEN_CDM_R7"

### Run in development
Open R in the package folder, or open project in Rstudio. 

```{r, eval=FALSE}
# load package
devtools::load_all(".")

# set configuration 
Sys.setenv(GOLEM_CONFIG_ACTIVE="<config_tag_in_golem-config.yml>")
# if this is not set, default configuration is no_connection environment
# Rstudio in sandbox is not reading the system environmental variables, force the envar as 
# Sys.setenv(BUCKET_SANDBOX_IVM="fg-production-sandbox-<n sandbox>_ivm")

# run shiny app
run_app()
)
```


## Deployment 

### Build
Clone repository, cd into main file and build giving as argument the github token:

```{bash, eval=FALSE}
git clone <CohortOperationsShinyApp_url>
cd <CohortOperationsShinyApp>
sudo docker build -t <docker_name> --build-arg GITHUB_PAT=<paste_token> .
```


### Run 




















