---
title: CohortOperationsShinyApp
author: Javier Gracia-Tabuenca
output:
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Intro

The goal of CohortOperationsShinyApp is to provide a graphical interface to create new cohorts based on cohorts imported from Atlas or from a text file. 

A cohort in this tool follows the epidemiological definition: "A cohort is a set of persons who satisfy one or more inclusion criteria for a duration of time" [[1]](https://ohdsi.github.io/TheBookOfOhdsi/Cohorts.html#what-is-a-cohort). 

This web app is a graphical interface to other R-packages which provide the underlying functionality. 
For programming use cases check these: 

- [FINNGEN/FinnGenTableTypes](https://github.com/FINNGEN/FinnGenTableTypes): Provides tools to create and operate cohortData tables. 
- [FINNGEN/CDMTools](https://github.com/FINNGEN/CDMTools): Provides connection to Atlas and Atlas' CDM databases. 

This project follows [Golem](https://engineering-shiny.org/golem.html) philosophy for the development of Shiny Apps, including modularisation, unit testing, shiny server testing, yalm configuration, and dockerisation. Moreover, this projects uses [renv](https://rstudio.github.io/renv/articles/renv.html) for pakage dependency management. 

## Development 
Before any of the following steps you need a github token to access the code in the private repositories. Set the environmental variable `GITHUB_PAT` with a token generated in [github.com/settings/tokens](https://github.com/settings/tokens). Token should include following permissions: "gist, repo, user, workflow". Following code helps you to do that: 

```{r, eval=FALSE}
# generate tocken 
usethis::create_github_token()

# copy and paste in 
Sys.setenv(GITHUB_PAT="<paste_token>")
```

### Install in laptop 
For development you can clone this repository and use `renv::` to install all the dependent packages. 

`renv::` automatically installs its self at the opening of the project. Then, run `renv::restore()` to install the dependent packages. 

### Install in Sandbox 
Github's zip can be uploaded to sandbox. However, dependencies cant be installed into the internetless sandbox IVM. 

The dependencies can be collected before in an machine with internet connection. 
Then zipped and uploaded to sandbox. Follow `renv::` instruction for a such situation: 

>First open the R package using RStudio on a computer with internet (note: this should have the same operating system and R version). Specify the folder where your packages are stored by setting the RENV_PATHS_CACHE location (run `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))`). Then run `renv::restore()` in the console. Manually move the study package to the environment without internet (this now includes all required R packages), activate the current project with renv::activate() and again run `Sys.setenv("RENV_PATHS_CACHE"=paste0(getwd(),"/renv/cache"))` followed by `renv::restore()` in the console.

Alternatively, you can find a pre-build zip for sandbox for some releases in the github repository. 

### Configure 
Configuration for your development environment can be set in `inst/golem-config.yml`. 
Currently this file includes three environments: 

- `no_connection`: To work with no connection to an Atlas instance. 
- `atlas-development`: To work with Atlas install in an ivm in the atlas-development project in GCP.  
- `sandbox`: To work in sandbox. This environment needs of an additional environmental variable `GCP_BILLING_PROJECT_ID` with bigquery's billing project name (eg. "fg-production-sandbox-4") 

### Run 

```{r, eval=FALSE}
# load package
devtools::load_all(".")

# set configuration 
Sys.setenv(GOLEM_CONFIG_ACTIVE="<config_tag_in_golem-config.yml>")
#Sys.setenv(GCP_BILLING_PROJECT_ID="fg-production-sandbox-<your sandbox number>" ) # necesary if in sandbox

# run shiny app
run_app()
)
```


## Deployment 

### Configure

Get a token from github [github.com/settings/tokens](https://github.com/settings/tokens).
Is recommended to set the live of the token to one day.  

Set the token in the environmental variable `GITHUB_PAT`

```{bash, eval=FALSE}
export GITHUB_PAT="<paste_token>")
```
      
### Build
Clone repository, cd into main file and Build:

```{bash, eval=FALSE}
git clone <CohortOperationsShinyApp>
cd <CohortOperationsShinyApp>
sudo docker build .
```


### Run 




















